\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-09-04 forwards-and-backwards.tex
.

\begin{proof}
    First note that by Lemma~\ref{lem:eta_f_bound},
    $\IE[\langle 1, \eta^N_t \rangle] \le \IE[\langle 1, \eta^N_0 \rangle] e^{Ct}$
    for some $C$.
    Now, let $M^*_t(f) = \sup_{0 \le s \le t} M_t(f)$,
    and as before $[M(f)]_t$ the quadratic variation of $M_t(f)$,
    so that by the Burkholder-Davis-Gundy inequality, there is a $C'$ so that
    \begin{align*}
        \IE\left[ M^*_t(1) \right]
        &\le
        C' \IE\left[ \sqrt{[M^*(1)]_t} \right]
        \\ &\le
        C'\left( 1 + \IE\left[ [M^*(1)]_t \right] \right)
        \\ &=
        C'\left( 1 + \frac{\theta}{N} \IE\left[
            \int_0^t
            \int_{\IR^d} \left\{
                \gamma(x, \eta^N_s)
                \int_{\IR^d} r(y, \eta^N_s) q_\theta(x, dy)
                + \mu(x, \eta^N_s)
            \right\} \eta_s^N(dx)
            dt
            \right] \right)
        \\ &=
        C'\left( 1 + \IE\left[
            \int_0^t
            \int_{\IR^d} \left\{
                \frac{2\theta}{N} \gamma(x, \eta^N_s) r(x, \eta^N_s)
                + \frac{1}{N} B(x, \eta^N_s)
            \right\} \eta_s^N(dx)
            dt
            \right] \right) .
    \end{align*}
    Since $\gamma(x, \eta) r(x, \eta)$ and $B(x, \eta)$ are bounded,
    and $\theta/N \to \alpha < \infty$,
    combining constants and again using Gronwall's inequality, we get that
    \begin{align*}
        \IE\left[ M^*_t(1) \right]
        % &\le
        % C'\left( 1 + \IE\left[
        %     \int_0^t
        %     \int_{\IR^d} \left\{ C'' \right\} \eta_s^N(dx)
        %     dt
        %     \right] \right) \\
        &\le
        C' + C'' \IE[ \langle 1, \eta_0^N \rangle ] e^{tC} .
    \end{align*}
    Taking suprema and expectations on both sides of equation~\eqref{eqn:eta_f_mgale_decomp},
    then again using the fact that $\gamma(x, \eta) B_f(x, \eta) + F(x, \eta) \le C$,
    \begin{align*}
        \IE\left[\sup_{0 \le s \le t} \langle 1, \eta^N_s \rangle \right]
        &\le
        \IE[\langle 1, \eta^N_0 \rangle]
        + \IE\left[
            \sup_{0 \le s \le t}
            \int_0^t \int_{\IR^d} \left\{
                \gamma(x, \eta^N_s)
                B_f(x, \eta^N_s)
                + F(x, \eta^N_s)
            \right\} \eta^N_s(dx) ds
        \right]
        + \IE[M^*_t(1)] 
        \\
        &\le
        \IE[\langle 1, \eta^N_0 \rangle]
        + C \IE\left[
            \int_0^t \sup_{0 \le s \le t} \langle 1, \eta^N_s \rangle ds
        \right]
        + C' + C'' \IE[ \langle 1, \eta_0^N \rangle ] e^{tC} .
    \end{align*}
    Once again applying Gronwall's inequality,
    \begin{align*}
        \IE\left[\sup_{0 \le s \le t} \langle 1, \eta^N_s \rangle \right]
        &\le
        C'' 
        \left(\IE[\langle 1, \eta^N_0 \rangle] + C'\right)
        \IE[ \langle 1, \eta_0^N \rangle ] e^{2tC} .
    \end{align*}
    For any $T$,
    the inequality on the right is bounded above by a constant $C(T)$ independent of $N$.
    As a result, for any $K > 0$,
    \begin{equation*}
    \limsup_{N \to \infty}
        \IP\left[ \sup_{0 \le s \le T} \langle 1, \eta^{N}_{s} \rangle \geq K \right]
        \leq
        \frac{C(T)}{K}.
    \end{equation*}
\end{proof}

@@

\begin{theorem} \label{TeoremOneStepConvergence}
    Let $(\eta_t)_{t \geq 0}$ be the scaled particle system with $r(x, \eta) \equiv \gamma(x, \eta) \equiv 1$.
    Suppose $F(x)$ satisfies:
    \begin{enumerate}
        \item $F(x)1_{x>0}$  is bounded above.
        \item $|F'(x)1_{x>0}|$ is bounded,
        \item and $||F''(x)||_\infty < \infty$.
    \end{enumerate}
    We will also suppose $\theta$ is big enough that $1- F(x)1_{x>0}/\theta$ is non-negative.

    We will assume $\eta_0 := \eta_0(\epsilon,N,\theta)$
    is a sequence of initial conditions for $(\eta_t)_{t \geq 0}$
    such that there exists a finite constant $C>0$ independent of $\epsilon$, $N$, and $\theta$
    such that
    \begin{enumerate}
        \item $\langle 1 , \eta_0 \rangle < C $
        \item  $\| \rho_\epsilon * \eta_0(x) \|_\infty < C$ for all $x \in \IR^d$
        \item $|\rho_\epsilon * \eta_0(x) - \rho_\epsilon * \eta_0(y)| \leq C \| x - y \|$ for all $x,y \in \IR^d$, and
        \item $\rho_\epsilon * \eta_0(x) dx$ converges weakly to $u_0(x)dx$ for some $u_0(x) \in \mathcal{C}^2(\IR^d)$.
    \end{enumerate}
    Furthermore, suppose that
    \begin{equation}
        \frac{\theta}{\epsilon^d N} + \frac{1}{\theta \epsilon^{2}} \to 0 . \label{ConditionsForConvergence}
    \end{equation}
    Fix $T>0$. Then, the measure $((\rho_\epsilon *\eta_t)(x) dx)_{t \geq 0}$ converges weakly in $D([0,T], \measures)$ to a measure $(\Xi_t(dx))_{t \geq 0}$. Moreover, up to a modification on a set of measure zero, $\Xi_t(x) = u_t(x)dx$, where $u_t(x)$ is a weak solution to:
    \[ \partial_t u = \frac{1}{2}\Delta u + u F(u) \]
    with initial condition $u_0$. That is, for every smooth function with compact support $f$ and $t>0$, we have:
    \begin{equation*}
     \int \partial_t  u_t(x) f(x) dx = \int u_t(x)\frac{1}{2} \Delta f(x) dx + \int f(x) u_t(x) F(u_t(x)) dx.
    \end{equation*}
\end{theorem}

@@

Since we're assuming that $\gamma$ and $r$ are identically 1,
the death rate is $1 + F(x, \eta) / \theta$.
For reference,
the martingale problem (REF) for $\eta_t$ gives that for any smooth function $\phi$,
\begin{align}
&
    \langle \phi(x), \eta_{t}(dx) \rangle
    -
    \langle \phi(x), \eta_{0}(dx)\rangle
    =
    M_t(\phi)
    +
    \int_{0}^{t} \langle \theta \int_{\IR^d}
            \left( \phi(y) -\phi(x) \right)
        q_\theta(x,dy), \eta_s(dx)
    \rangle ds
\nonumber \\
&
    {} +
    \int_0^t \langle
        F(x, \eta_s) \phi(x),
        \eta_s(dx)
    \rangle ds , \label{mgp1}
\end{align}
where $M_t(\phi)$ is a martingale with quadratic variation:
\begin{align}
[ M(\phi) ]_t
    &:=
    \frac{\theta}{N}
    \int_{0}^{t} \bigg\{
        \left\langle \int_{\IR^d} \phi^2(y) q_\theta(x,dy), \eta_s(dx) \right\rangle
    \nonumber\\
    & {} +
    \left\langle \left(1 + \frac{F(x, \eta_s)}{\theta } \right)
        \phi^2(x), \eta_s(dx)) \right\rangle 
    \bigg\} ds.  \label{mgp2}
\end{align}

@@

We will require also a second moment bound on the total mass.

\begin{proposition}
Under the assumptions of Theorem~\ref{TeoremOneStepConvergence} there is a finite constant $C$ (independent of $\epsilon, N, \theta$) such that, for all $t \in [0,T]$ and for big enough $N$ and $\theta$, $\IE[ \langle 1, \eta_t \rangle^2] \leq C$.\label{Prop:SecondMomentMass}
\end{proposition}

\begin{proof}[of Proposition~\ref{Prop:SecondMomentMass}]
    % Referring to section~\ref{sec:heuristics} with $f=1$ and $G(m) = m^2$,
    When the population is in state $\eta$,
    individuals of mass $1/N$ are born at rate $\theta \langle 1, \eta \rangle$
    and die at rate $\theta \langle 1 + F/\theta, \eta \rangle$.
    Similar to the arguments in section~\ref{sec:heuristics} with $f=1$ and $G(m) = m^2$,
    since $N G(m \pm 1/N) - G(m) = \pm 2m + 1/N$,
    \begin{align*}
        \IE[\langle 1, \eta_t \rangle^2]
        &=
        \IE[\langle 1, \eta_0 \rangle^2]
        +
        \IE\left[
        \int_0^t
        \left\{
            \theta \langle 1, \eta_s \rangle
            \left( 2 \langle 1, \eta_s \rangle + 1/N \right)
            +
            \theta \langle 1 + F/\theta, \eta_s \rangle
            \left( - 2 \langle 1, \eta_s \rangle + 1/N \right)
        \right\}
        ds
        \right]
        \\
        &=
        \IE[\langle 1, \eta_0 \rangle^2]
        +
        \IE\left[
        \int_0^t \left\{
            \frac{2\theta}{N} \langle 1, \eta_s \rangle ds
            -
            2 \langle F, \eta_s \rangle \left(\langle 1, \eta_s \rangle - \frac{1}{N} \right)
        \right\} ds \right].
    \end{align*}
    Since $F(m)1_{m > 0}$ is bounded above and $\theta/N$ converges,
    we have for an appropriate $C$ that
    \begin{align*}
        \IE[\langle 1, \eta_t \rangle^2]
        \le
        C + C \int_0^t \IE[\langle 1, \eta_s \rangle] ds
        + C \int_0^t \IE[\langle 1, \eta_s \rangle^2] ds .
    \end{align*}
    The result follows from Lemma~\ref{Lemma:UniformlyBoundedMass} and Gronwall's inequality.
\end{proof}

@@

\begin{lemma} \label{PsiBoundHS}
    Fix $t>0$,
    let $(N(s))_{s \ge 0}$ be a rate 1 Poisson process,
    and let $T(t) = N(\theta t) / \theta$.
    Then
    \[
        \psi^{\epsilon,x}_t(y)
        =
        \IE\left[ p_{\epsilon^2+T(t)}(x,y)\right],
    \]
    and moreover, there is a $C$ independent of $\epsilon$ or $t$ such that
    \[
        \| \psi^{\epsilon,x}_t \|_\infty
        \leq
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        \left(
            1 + \exp\left(-\frac{\epsilon^2 \theta}{2}\right)
        \right) .
    \]
\end{lemma}

@@

    For the second claim, first define $\tau(t) = T(t)-t$.
    % which has mean $0$ and variance $t/\theta$.
    % so by Chebyshev's inequality,
    % \begin{equation}
    %     \IP\left\{ \tau(t) < - \frac{\epsilon^2 + t}{2} \right\}
    %     \leq
    %     \IP\left\{  \frac{|T(t)-t|}{\epsilon^2+t} > \frac{1}{2} \right\}
    %     \leq
    %     \frac{4t}{\theta(\epsilon^2+t)^2} \leq \frac{2}{\epsilon^2 \theta}.
    %     \label{eq:IneqConcentrationT}
    % \end{equation}
    Since if $\tau(t) \ge -(\epsilon^2 + t)/2$,
    then $1/(\epsilon^2 + T(t)) \le 2/(\epsilon^2 + t)$,
    while $\epsilon^2 + T(t) \ge \epsilon^2$ always,
    then partitioning over
    $\{ \tau(t) \ge -(\epsilon^2 + t)/2 \}$ and its complement,
    \begin{align}
    \| \psi^{\epsilon,x} \|_\infty
    &=
        \IE\left[ 
            \frac{1}{2 \pi (\epsilon^2+T(t))^{d/2}}
        \right] \nonumber
    \\ & \le
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        + 
        \frac{C}{\epsilon^d}
        \IP\left\{
            \tau(t) < - (\epsilon^2 + t) / 2
        \right\} \label{eqn:psi_infty_bound}
    \end{align}
    Now, observe that since $\IE[e^{-N(\theta t)}] = \exp(-\theta t (1 - e^{-1}))$,
    then by Markov's inequality,
    \begin{align*}
        \IP\left\{
            \tau(t) < - \frac{\epsilon^2 + t}{2}
        \right\}
        &=
        \IP\left\{
            e^{-N(\theta t)} > e^{-\theta (t - \epsilon^2) / 2}
        \right\}
        \\&\le
        \frac{
            \exp(- \theta t(1 - e^{-1}))
        }{
            \exp(- \theta (t - \epsilon^2) / 2)
        }
        \\&=
        \exp\left\{ - \delta \theta t - \frac{\theta \epsilon^2}{2} \right\} ,
    \end{align*}
    where $\delta = 1/2 - e^{-1} > 0$.
    The second term in \eqref{eqn:psi_infty_bound} is therefore bounded by
    \[
        C \left(1 + \frac{t}{\epsilon^2}\right)^{d/2} e^{-\delta \theta t}
        \frac{1}{(\epsilon^2 + t)^{d/2}} e^{-\epsilon^2 \theta / 2} .
    \]
    Now, $e^{-\delta \theta t} (1 + t/\epsilon^2)^{d/2}$ is decreasing in $t$
    for all $t$ if $\epsilon^2 \theta$ is sufficiently large:
    this holds because its derivative is
    is
    \[
        \left( \frac{d}{2 \epsilon^2} - \left(1 + \frac{t}{\epsilon^2}\right) \delta \theta \right)
        \left( 1 + \frac{t}{\epsilon^2} \right)^{d/2 - 1} e^{-\delta \theta t} ,
    \]
    which is negative if $\theta(\epsilon^2 + t) > d/2 \delta$.
    Therfore, the second term in \eqref{eqn:psi_infty_bound} 
    is bounded by $C e^{-\epsilon^2 \theta / 2} / (\epsilon^2 + t)^{d/2}$
    (for a different constant $C$).
    Substituting this into \eqref{eqn:psi_infty_bound} obtains the result.
\end{proof}

We will need a further comparison of $\psi_t^{\epsilon,x}$ to the heat semigroup,
and for this we will first need the following fact about large deviations
of a Poisson process.

@@

\begin{lemma}
    \label{lem:psi_approx_p}
    In the notation of Lemma~\ref{PsiBoundHS},
    and with $A$ as in Lemma~\ref{lem:poisson_ld},
    \begin{align*}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \le
        \frac{C}{\epsilon^2 \theta}^{d/2}
        p_{2(A+1)(\epsilon^2+t)}(x, y)
        +
        C \exp(- \epsilon^2 \theta / 2)
        +
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        \IP\left\{
            \tau(t) > A(\epsilon^2 + t)
        \right\} .
    \end{align*}
\end{lemma}

\begin{proof}[of Lemma~\ref{lem:psi_approx_p}]
    We will partition into three events according to the value of $\tau(t)$:
    let $A_1 = \{ \tau(t) < - (\epsilon^2 + t)/2 \}$,
    let $A_2 = \{ \tau(t) > A(\epsilon^2 + t) \}$,
    and $A_3$ the remaining event, $\{ - (\epsilon^2 + t)/2 \le \tau(t) \le A(\epsilon^2 + t) \}$.
    Then,
    \begin{align*}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        &=
        \left|
        \IE\left[
            p_{\epsilon^2 + t + \tau(t)}(x, y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right]
        \right|
        \\&\le
        \IE\left[
            (1_{A_1} + 1_{A_2} + 1_{A_3})
            \left|
            p_{\epsilon^2 + t + \tau(t)}(x, y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \right] .
    \end{align*}

@@

\begin{lemma}
Let $s>0, x, y, z\in \IR^d$. The following estimate holds
\[ |p_s(x,z) - p_s(y,z)| \leq \frac{C|x-y|}{\sqrt{s}} \left(p_{2s}(x,z) + p_{2s}(y,z)\right) \]
where the constant $C$ does not depend on $x,y,z$ or $s$.\label{Lemma:ContinuityHS}
\end{lemma}
Expanding the difference of two squares we can rewrite,
&\exp\left(-\frac{(y-z)^2}{2 s} \right)-\exp\left(-\frac{(x-z)^2}{2 s} \right) \\ &= \left(\exp\left(-\frac{(y-z)^2}{4 s} \right)-\exp\left(-\frac{(x-z)^2}{4 s} \right) \right)\left(\exp\left(-\frac{(y-z)^2}{4 s} \right)+\exp\left(-\frac{(x-z)^2}{4 s} \right) \right)
by using the Intermediate Value Theorem we can write the right hand side as,
\[ (y-x) \left( - \frac{2(w-z)}{4s} \exp\left(-\frac{(w-z)^2}{4s}\right)\right) 2^{d/2}\left( p_{2s}(y,z)+p_{2s}(x,z) \right)\]
for some $w \in [y,z]$. Using the fact that $x e^{-x^2}$ is uniformly bounded we can bounded the first bracket in the last equation by $C/\sqrt{s}$ and the result follows.

>>>>

\begin{proof}
    First note that by Lemma~\ref{lem:eta_f_bound},
    $\IE[\langle 1, \eta^N_t \rangle] \le \IE[\langle 1, \eta^N_0 \rangle] e^{Ct}$
    for some $C$.
    Now, let $M^*_t(f) = \sup_{0 \le s \le t} M_t(f)$,
    and as before $[M(f)]_t$ the quadratic variation of $M_t(f)$,
    so that by the Burkholder-Davis-Gundy inequality, there is a $C'$ so that
    \begin{align*}
        \IE\left[ M^*_t(1) \right]
        &\le
        C' \IE\left[ \sqrt{[M(1)]_t} \right]
        \\ &\le
        C'\left( 1 + \IE\left[ [M(1)]_t \right] \right)
        \\ &=
        C'\left( 1 + \frac{\theta}{N} \IE\left[
            \int_0^t
            \int_{\IR^d} \left\{
                \gamma(x, \eta^N_s)
                \int_{\IR^d} r(y, \eta^N_s) q_\theta(x, dy)
                + \mu(x, \eta^N_s)
            \right\} \eta_s^N(dx)
            dt
            \right] \right)
        \\ &=
        C'\left( 1 + \IE\left[
            \int_0^t
            \int_{\IR^d} \left\{
                \frac{2\theta}{N} \gamma(x, \eta^N_s) r(x, \eta^N_s)
                + \frac{1}{N} B(x, \eta^N_s)
            \right\} \eta_s^N(dx)
            dt
            \right] \right) .
    \end{align*}
    Since $\gamma(x, \eta) r(x, \eta)$ and $B(x, \eta)$ are bounded,
    and $\theta/N \to \alpha < \infty$,
    combining constants and again using Gronwall's inequality, we get that
    \begin{align*}
        \IE\left[ M^*_t(1) \right]
        % &\le
        % C'\left( 1 + \IE\left[
        %     \int_0^t
        %     \int_{\IR^d} \left\{ C'' \right\} \eta_s^N(dx)
        %     dt
        %     \right] \right) \\
        &\le
        C' + C'' \IE[ \langle 1, \eta_0^N \rangle ] e^{tC} .
    \end{align*}
    Taking suprema and expectations on both sides of equation~\eqref{eqn:eta_f_mgale_decomp},
    then again using the fact that $\gamma(x, \eta) B_f(x, \eta) + F(x, \eta) \le C$,
    \begin{align*}
        \IE\left[\sup_{0 \le s \le t} \langle 1, \eta^N_s \rangle \right]
        &\le
        \IE[\langle 1, \eta^N_0 \rangle]
        + \IE\left[
            \sup_{0 \le s \le t}
            \int_0^t \int_{\IR^d} \left\{
                \gamma(x, \eta^N_s)
                B_f(x, \eta^N_s)
                + F(x, \eta^N_s)
            \right\} \eta^N_s(dx) ds
        \right]
        + \IE[M^*_t(1)] 
        \\
        &\le
        \IE[\langle 1, \eta^N_0 \rangle]
        + C \IE\left[
            \int_0^t \sup_{0 \le s \le t} \langle 1, \eta^N_s \rangle ds
        \right]
        + C' + C'' \IE[ \langle 1, \eta_0^N \rangle ] e^{tC} .
    \end{align*}
    Once again applying Gronwall's inequality,
    \begin{align*}
        \IE\left[\sup_{0 \le s \le t} \langle 1, \eta^N_s \rangle \right]
        &\le
        C'' 
        \left(\IE[\langle 1, \eta^N_0 \rangle] + C'\right)
        \IE[ \langle 1, \eta_0^N \rangle ] e^{2tC} .
    \end{align*}
    For any $T$,
    the inequality on the right is bounded above by a constant $C(T)$ independent of $N$.
    As a result, for any $K > 0$,
    \begin{equation*}
    \limsup_{N \to \infty}
        \IP\left[ \sup_{0 \le s \le T} \langle 1, \eta^{N}_{s} \rangle \geq K \right]
        \leq
        \frac{C(T)}{K}.
    \end{equation*}
\end{proof}

@@

\begin{theorem} \label{TeoremOneStepConvergence}
    Let $(\eta_t)_{t \geq 0}$ be the scaled particle system with $r(x, \eta) \equiv \gamma(x, \eta) \equiv 1$.
    Suppose $F(x)$ satisfies for $x \ge 0$:
    \begin{enumerate}
        \item $F(x)$  is bounded above.
        \item $|F'(x)|$ is bounded,
        \item and $||F''(x)||_\infty < \infty$.
    \end{enumerate}
    We will also suppose $\theta$ is big enough that $1- F(x)/\theta$ is non-negative.

    We will assume $\eta_0 := \eta_0(\epsilon,N,\theta)$
    is a sequence of initial conditions for $(\eta_t)_{t \geq 0}$
    such that there exists a finite constant $C>0$ independent of $\epsilon$, $N$, and $\theta$
    such that
    \begin{enumerate}
        \item[(i)] $\langle 1 , \eta_0 \rangle < C $
        \item[(ii)]  $\| \rho_\epsilon * \eta_0(x) \|_\infty < C$ for all $x \in \IR^d$
        \item[(iii)] $|\rho_\epsilon * \eta_0(x) - \rho_\epsilon * \eta_0(y)| \leq C \| x - y \|$ for all $x,y \in \IR^d$, and
        \item[(iv)] $\rho_\epsilon * \eta_0(x) dx$ converges weakly to $u_0(x)dx$ for some $u_0(x) \in \mathcal{C}^2(\IR^d)$.
    \end{enumerate}
    Furthermore, suppose that
    \begin{equation}
        \frac{\theta}{\epsilon^d N} + \frac{1}{\theta \epsilon^{2}} \to 0 . \label{ConditionsForConvergence}
    \end{equation}
    Fix $T>0$. Then, the measure $((\rho_\epsilon *\eta_t)(x) dx)_{t \geq 0}$ converges weakly in $D([0,T], \measures)$ to a measure $(\Xi_t(dx))_{t \geq 0}$. Moreover, up to a modification on a set of measure zero, $\Xi_t(x) = u_t(x)dx$, where $u_t(x)$ is a weak solution to:
    \[ \partial_t u = \frac{1}{2}\Delta u + u F(u) \]
    with initial condition $u_0$. That is, for every smooth function with compact support $f$ and $t>0$, we have:
    \begin{equation*}
     \int \partial_t  u_t(x) f(x) dx = \int u_t(x)\frac{1}{2} \Delta f(x) dx + \int f(x) u_t(x) F(u_t(x)) dx.
    \end{equation*}
\end{theorem}

@@

Since we're assuming that $\gamma$ and $r$ are identically 1,
the death rate is $1 + F(x, \eta) / \theta$.
For reference,
the martingale problem (Definition \ref{defn:mgale_construction})
for $\eta_t$ gives that for any smooth function $\phi$,
\begin{align} \label{mgp1}
    \langle \phi(x), \eta_{t}(dx) \rangle
    -
    \langle \phi(x), \eta_{0}(dx)\rangle
    &= \nonumber
    M_t(\phi)
    +
    \int_{0}^{t} \langle \theta \int_{\IR^d}
            \left( \phi(y) -\phi(x) \right)
        q_\theta(x,dy), \eta_s(dx)
    \rangle ds
    \\ & {}
    +
    \int_0^t \langle
        F(x, \eta_s) \phi(x),
        \eta_s(dx)
    \rangle ds ,
\end{align}
where $M_t(\phi)$ is a martingale with quadratic variation:
\begin{align} \label{mgp2}
    \begin{split}
[ M(\phi) ]_t
    &:=
    \frac{\theta}{N}
    \int_{0}^{t} \bigg\{
        \left\langle \int_{\IR^d} \phi^2(y) q_\theta(x,dy), \eta_s(dx) \right\rangle
    \\
    & \qquad \qquad {} +
    \left\langle \left(1 + \frac{F(x, \eta_s)}{\theta } \right)
        \phi^2(x), \eta_s(dx)) \right\rangle 
    \bigg\} ds. 
    \end{split}
\end{align}

Several times below we will need to bound second moments,
and for this purpose it will be useful to record that \eqref{mgp1} and \eqref{mgp2}
imply that the following is a martingale:
\begin{align} \label{mgp3}
    \nonumber
    \langle \phi, \eta_t \rangle^2
    -
    & \langle \phi, \eta_0 \rangle^2
    -
    \int_0^t
        2 \langle \phi, \eta_s \rangle
        \left\langle
            \theta \int_{\IR^d} (\phi(y) - \phi(x)) q_\theta(x, dy)
            + F(x, \eta_s) \phi(x),
            \eta_s(dx)
        \right\rangle
    ds
    \\ & {}
    -
    \frac{\theta}{N}
    \int_0^t
        \left\langle
            \int_{\IR^d} \phi^2(y) q_\theta(x, dy)
            +
            \left(1 + \frac{F(x, \eta_s)}{\theta}\right) \phi^2(x),
            \eta_s(dx)
        \right\rangle
    ds .
\end{align}

@@

\begin{lemma}
        \label{lem:SecondMomentMass}
    Under the assumptions of Theorem~\ref{TeoremOneStepConvergence}
    there is a finite constant $C$ (independent of $\epsilon, N, \theta$) such that,
    for all $t \in [0,T]$ and for big enough $N$ and $\theta$,
    $\IE[ \langle 1, \eta_t \rangle^2] \leq C$.
\end{lemma}

\begin{proof}[of Lemma~\ref{lem:SecondMomentMass}]
    Using equation~\eqref{mgp3} with $\phi = 1$,
    \begin{align*}
        \IE[\langle 1, \eta_t \rangle^2]
        &=
        \IE[\langle 1, \eta_0 \rangle^2]
        +
        \IE\left[
        \int_0^t
            2 \langle 1, \eta_s \rangle
            \langle F(x, \eta_s), \eta_s(dx) \rangle
            +
            \frac{\theta}{N} \langle 2 + F(x, \eta_s)/\theta, \eta_s(dx) \rangle
        ds
        \right] .
    \end{align*}
    Since $F(m)1_{m > 0}$ is bounded above and $\theta/N$ converges,
    we have for an appropriate $C$ that
    \begin{align*}
        \IE[\langle 1, \eta_t \rangle^2]
        \le
        C + C \int_0^t \IE[\langle 1, \eta_s \rangle] ds
        + C \int_0^t \IE[\langle 1, \eta_s \rangle^2] ds .
    \end{align*}
    The result follows from Lemma~\ref{Lemma:UniformlyBoundedMass} and Gronwall's inequality.
\end{proof}

@@

\begin{lemma} \label{PsiBoundHS}
    Fix $t>0$,
    let $(N(s))_{s \ge 0}$ be a rate 1 Poisson process,
    and let $T(t) = N(\theta t) / \theta$.
    Then
    \[
        \psi^{\epsilon,x}_t(y)
        =
        \IE\left[ p_{\epsilon^2+T(t)}(x,y)\right],
    \]
    and moreover, there is a $C$ independent of $\epsilon$ or $t$ such that
    \[
        \| \psi^{\epsilon,x}_t \|_\infty
        \leq
        \frac{C}{(\epsilon^2 + t)^{d/2}} .
    \]
\end{lemma}

@@

    For the second claim, first define $\tau(t) = T(t)-t$.
    % which has mean $0$ and variance $t/\theta$.
    % so by Chebyshev's inequality,
    % \begin{equation}
    %     \IP\left\{ \tau(t) < - \frac{\epsilon^2 + t}{2} \right\}
    %     \leq
    %     \IP\left\{  \frac{|T(t)-t|}{\epsilon^2+t} > \frac{1}{2} \right\}
    %     \leq
    %     \frac{4t}{\theta(\epsilon^2+t)^2} \leq \frac{2}{\epsilon^2 \theta}.
    %     \label{eq:IneqConcentrationT}
    % \end{equation}
    Since if $\tau(t) \ge -(\epsilon^2 + t)/2$,
    then $1/(\epsilon^2 + T(t)) \le 2/(\epsilon^2 + t)$,
    while $\epsilon^2 + T(t) \ge \epsilon^2$ always,
    then partitioning over
    $\{ \tau(t) \ge -(\epsilon^2 + t)/2 \}$ and its complement,
    \begin{align}
    \| \psi^{\epsilon,x} \|_\infty
    &=
        \IE\left[ 
            \frac{1}{2 \pi (\epsilon^2+T(t))^{d/2}}
        \right] \nonumber
    \\ & \le
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        + 
        \frac{C}{\epsilon^d}
        \IP\left\{
            \tau(t) < - (\epsilon^2 + t) / 2
        \right\} \label{eqn:psi_infty_bound}
    \end{align}
    Now, observe that since $\IE[e^{-N(\theta t)}] = \exp(-\theta t (1 - e^{-1}))$,
    then by Markov's inequality,
    \begin{align*}
        \IP\left\{
            \tau(t) < - \frac{\epsilon^2 + t}{2}
        \right\}
        &=
        \IP\left\{
            e^{-N(\theta t)} > e^{-\theta (t - \epsilon^2) / 2}
        \right\}
        \\&\le
        \frac{
            \exp(- \theta t(1 - e^{-1}))
        }{
            \exp(- \theta (t - \epsilon^2) / 2)
        }
        \\&=
        \exp\left\{ - \delta \theta t - \frac{\theta \epsilon^2}{2} \right\} ,
    \end{align*}
    where $\delta = 1/2 - e^{-1} > 0$.
    The second term in \eqref{eqn:psi_infty_bound} is therefore bounded by
    \[
        C \left(1 + \frac{t}{\epsilon^2}\right)^{d/2} e^{-\delta \theta t - \epsilon^2 \theta / 2}
        \frac{1}{(\epsilon^2 + t)^{d/2}} .
    \]
    Now, $e^{-\delta \theta t - \epsilon^2 \theta / 2} (1 + t/\epsilon^2)^{d/2}$ is decreasing in $t$
    for all $t$ if $\epsilon^2 \theta$ is sufficiently large:
    this holds because its derivative is
    is
    \[
        \left( \frac{d}{2 \epsilon^2} - \left(1 + \frac{t}{\epsilon^2}\right) \delta \theta \right)
        \left( 1 + \frac{t}{\epsilon^2} \right)^{d/2 - 1} e^{-\delta \theta t - \epsilon^2 \theta / 2} ,
    \]
    which is negative if $\theta(\epsilon^2 + t) > d/2 \delta$.
    Therfore, the second term in \eqref{eqn:psi_infty_bound} 
    is bounded by $C / (\epsilon^2 + t)^{d/2}$
    (for a different constant $C$).
    Substituting this into \eqref{eqn:psi_infty_bound} obtains the result.
\end{proof}

We will need further comparisons of $\psi_t^{\epsilon,x}$ to the heat semigroup,
and for this we will first need the following fact about large deviations
of a Poisson process.

@@

\begin{lemma}
    \label{Lemma:BoundPsiHS2}
    In the notation of Lemma~\ref{PsiBoundHS},
    and with $A$ as in Lemma~\ref{lem:poisson_ld},
    \begin{align*}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \le
        \frac{C}{\epsilon^2 \theta}^{d/2}
        p_{2(A+1)(\epsilon^2+t)}(x, y)
        +
        C \exp(- \epsilon^2 \theta / 2)
        +
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        \IP\left\{
            \tau(t) > A(\epsilon^2 + t)
        \right\} .
    \end{align*}
\end{lemma}

\begin{proof}[of Lemma~\ref{Lemma:BoundPsiHS2}]
    The three terms on the right hand side of the Lemma
    correspond to bound obtained by partitioning
    into three events according to the value of $\tau(t)$:
    let $A_1 = \{ \tau(t) < - (\epsilon^2 + t)/2 \}$,
    let $A_2 = \{ \tau(t) > A(\epsilon^2 + t) \}$,
    and $A_3$ the remaining event, $\{ - (\epsilon^2 + t)/2 \le \tau(t) \le A(\epsilon^2 + t) \}$.
    Then,
    \begin{align*}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        &=
        \left|
        \IE\left[
            p_{\epsilon^2 + t + \tau(t)}(x, y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right]
        \right|
        \\&\le
        \IE\left[
            (1_{A_1} + 1_{A_2} + 1_{A_3})
            \left|
            p_{\epsilon^2 + t + \tau(t)}(x, y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \right] .
    \end{align*}

@@

\begin{lemma}
    \label{Lemma:ContinuityHS}
Let $s>0, x, y, z\in \IR^d$. The following estimate holds:
\[
    |p_s(x,z) - p_s(y,z)|
    \leq
    \frac{C|x-y|}{\sqrt{s}} \left(p_{2s}(x,z) + p_{2s}(y,z)\right) ,
\]
where the constant $C$ does not depend on $x,y,z$ or $s$.
\end{lemma}
